Modelling the Real Time State of a Public Transport Road Network

### 2017-04-26 realtime-filter/realtime-filter.Rnw
fix up things

\begin{equation}
  \label{eq:l1}
  \log L_1 (\bX_k^{(i)} | \bY_k) = - \log(2\pi\sigma_y) - \frac{\|g(h(d_k^{(i)})|\bY_k)\|^2}{2\sigma_y^2}
\end{equation}

@@

\begin{equation}
  \label{eq:l1b}
  \log L_1(\bX_k^{(i)} | \bY_k)=
  \begin{cases}
    - \log(\pi \varepsilon_S^2) & \bY_k \in B_s \\
    - \infty & \text{otherwise}
  \end{cases}
\end{equation}

@@

\begin{equation}
  \label{eq:l1c}
  \log L_1 (\bX_k^{(i)} | \bY_k) =
  \begin{cases}
    - \log(A) & \bY_k \in B_I^1 \cap B_I^2 \\
    - \infty & \text{otherwise}
  \end{cases}
\end{equation}
where $A = 2\varepsilon_I^2 \left( \cos^{-1}(\frac{1}{2}) - \frac{\sqrt{3}}{4} \right)$,
the area of $B_I^1\cap B_I^2$.

@@

\begin{enumerate}
\item 
  If both $T^a_j$ and $T^d_j$ are NULL, 
  then there is no change to the likelihood;
  $ L_2  (\bX_k^{(i)} | \bT_k) = 1 $.

\item
  If departure time, $T^d_j$, is available only, for stop $j$, then we
  know the bus has at least departed stop $j$; therefore,
  \begin{equation}
    \label{eq:l2}
    L_2 (\bX_k^{(i)} | \bT_k) =
      \begin{cases}
        f_N(D_j^{(i)} | T_j^d, \sigma_d^2) & s_k^{(i)} \geq j \cap D_j^{(i)} \neq \text{NULL} \\
        0 & \text{otherwise},
      \end{cases}
  \end{equation}
  where $f_N(x | \mu, \sigma^2)$ is the probability density of a normal distribution
  with mean $\mu$ and variance $\sigma^2$.

\item
  If arrival time, $T^a_j$, is available only, then similarly
  \begin{equation}
    \label{eq:l2b}
    L_2  (\bX_k^{(i)} | \bT_k) =
    \begin{cases}
      f_N(A_j^{(i)} | T_j^a, \sigma_d^2) & s_k^{(i)} \geq j \\
      0 & \text{otherwise}.
    \end{cases}
  \end{equation}
  Note that this does not assume the particle hasn't yet left the stop;
  the departure time could simply be missing data.

\item
  If both arrival time and departure time are available, then
  \begin{equation}
    \label{eq:l2c}
    L_2  (\bX_k^{(i)} | \bT_k) =
    \begin{cases}
      \begin{array}{l}
        f_N(A_j^{(i)} | T_j^a, \sigma_d^2) f_E(D_k^{(i)} - A_k^{(i)} | T_k^\delta) \quad\\
        \hfill s_k^{(i)} \geq j \cap D_j^{(i)} \neq \text{NULL} \\[0.5em]
        0 \qquad\qquad \text{otherwise},
      \end{array}
    \end{cases}
  \end{equation}
  where $T_k^\delta = T_k^d - T_k^a$ is the observed dwell time at the stop,
  and $f_E(x | \lambda)$ is the probability density of an exponential distribution 
  with rate $\lambda$.

@@

We now determine the likelihood of each particle by combining $L_1$ and $L_2$;
\begin{equation}
  \label{eq:particle_likelihood}
  L (\bX_k^{(i)} | \bY_k, \bT_k) = L_1(\bX_k^{(i)}|\bY_k) L_2(\bX_k^{(i)} | \bT_k).
\end{equation}
This is then used to compute the weight of each particle,
\begin{equation}
  \label{eq:particle_weights}
  w_i = \frac{L (\bX_k^{(i)} | \bY_k, \bT_k)}{\sum_{j=1}^N L (\bX_k^{(j)} | \bY_k, \bT_k)}.
\end{equation}

@@

In situations where multiple vehicles are in the same segment $\ell$, 
we can simply combine the estimates, using $M_\ell = $ number of vehicles in segment $\ell$ 
\cite{cn}.
\begin{equation}
  \label{eq:kf_data_estimates}
  \bar z_\ell = \frac{1}{M_\ell} \sum_{j = 1}^{M_\ell} \bar v_j \quad\text{and}\quad
  \hat \bR_{\ell,\ell}^2 = \frac{1}{M_\ell} \sum_{j=1}^{M_\ell} \left(\bar v_j^2 + \bar S_j^2\right) - \hat v_\ell^2
\end{equation}

>>>>

\begin{equation}
  \label{eq:l1}
  \log L_1 (\bY_k | \bX_k^{(i)}) = - \log(2\pi\sigma_y) - \frac{\|g(h(d_k^{(i)})|\bY_k)\|^2}{2\sigma_y^2}
\end{equation}

@@

\begin{equation}
  \label{eq:l1b}
  \log L_1(\bY_k | \bX_k^{(i)})=
  \begin{cases}
    - \log(\pi \varepsilon_S^2) & \bY_k \in B_s \\
    - \infty & \text{otherwise}
  \end{cases}
\end{equation}

@@

\begin{equation}
  \label{eq:l1c}
  \log L_1 (\bY_k | \bX_k^{(i)}) =
  \begin{cases}
    - \log(A) & \bY_k \in B_I^1 \cap B_I^2 \\
    - \infty & \text{otherwise}
  \end{cases}
\end{equation}
where $A = 2\varepsilon_I^2 \left( \cos^{-1}(\frac{1}{2}) - \frac{\sqrt{3}}{4} \right)$,
the area of $B_I^1\cap B_I^2$.

@@

\begin{enumerate}
\item 
  If both $T^a_j$ and $T^d_j$ are NULL, 
  then there is no change to the likelihood;
  $ L_2  (\bT_k | \bX_k^{(i)}) = 1 $; else

\item
  If $s_k^{(i)} \neq j$ (particle's stop index isn't the same as the stop time update's stop sequence),
  then we check particle is plausible given the data:
  \begin{equation}
    \label{eq:l2a}
    L_2 (\bT_k | \bX_k^{(i)}) =
    \mathbbm{1}_{s_k^{(i)} > j\ \cap\ A_{s_k}^{(i)} > T_j^a\ \cap\ A_{s_k}^{(i)} > T_j^d},
  \end{equation}
  (defining $x > y$ as TRUE if either $x$ or $y$ are NULL to simplify notation); else

\item
  If departure time, $T^d_j$, is available only, for stop $j$, then we
  know the bus has departed stop $j$; therefore,
  \begin{equation}
    \label{eq:l2}
    L_2 (\bT_k | \bX_k^{(i)}) =
        f_N(T_j^d | D_j^{(i)}, \sigma_d^2) \mathbbm{1}_{D_j^{(i)}}
  \end{equation}
  where $f_N(x | \mu, \sigma^2)$ is the probability density of a normal distribution
  with mean $\mu$ and variance $\sigma^2$,
  and $\mathbbm{1}_Z = 0$ if $Z$ is NULL, and $1$ otherwise; else

\item
  If arrival time, $T^a_j$, is available only, then similarly
  \begin{equation}
    \label{eq:l2b}
    L_2  (\bT_k | \bX_k^{(i)}) =
      f_N(T_j^a | A_j^{(i)}, \sigma_d^2)
  \end{equation}
  Note that this does not assume the particle hasn't yet left the stop;
  the departure time could simply be missing data; else

\item
  If both arrival time and departure time are available, then
  \begin{equation}
    \label{eq:l2c}
    L_2  (\bT_k | \bX_k^{(i)}) =
        f_N(T_j^a| A_j^{(i)}, \sigma_d^2) f_E(T_k^\delta | D_k^{(i)} - A_k^{(i)}) \mathbbm{1}_{D_j^{(i)}}
  \end{equation}
  where $T_k^\delta = T_k^d - T_k^a$ is the observed dwell time at the stop,
  and $f_E(x | \lambda)$ is the probability density of an exponential distribution 
  with rate $\lambda$.

@@

We now determine the likelihood of each particle by combining $L_1$ and $L_2$;
\begin{equation}
  \label{eq:particle_likelihood}
  L ([\bY_k, \bT_k] | \bX_k^{(i)}) = L_1(\bY_k|\bX_k^{(i)}) L_2(\bT_k | \bX_k^{(i)}).
\end{equation}
This is then used to compute the weight of each particle,
\begin{equation}
  \label{eq:particle_weights}
  w_i = \frac{L ([\bY_k, \bT_k] | \bX_k^{(i)})}{\sum_{j=1}^N   L ([\bY_k, \bT_k] | \bX_k^{(j)})}.
\end{equation}

@@

In situations where multiple vehicles are in the same segment $\ell$, 
we can simply combine the estimates, using $M_\ell = $ number of vehicles in segment $\ell$ 
\cite{cn}.
\begin{equation}
  \label{eq:kf_data_estimates}
  \bar z_\ell = \frac{1}{M_\ell} \sum_{j = 1}^{M_\ell} \bar v_j \quad\text{and}\quad
  \hat \bR_{\ell,\ell}^2 = \frac{1}{M_\ell} \sum_{j=1}^{M_\ell} \left(\bar v_j^2 + \bar S_j^2\right) - \bar z_\ell^2
\end{equation}
